% !TEX root = main.tex
\section{Data Availability Statement}\label{data-availability-statement}

Implementation code, prompts, simulation transcripts, statistical analysis materials, and related artifacts are available at \url{https://github.com/SamuelDevdas/socialbehaviourregulation-SD}. Additional evaluation materials (including author qualitative review notes), if not present in the repository, may be requested from the corresponding author.

\section{Author Contributions}\label{author-contributions}

Conceptualization, Samuel Devdas, Dr.~Guang Lu, and Dr.~Alexandre de Spindler; Methodology, Samuel Devdas; Software, Samuel Devdas; Validation, Samuel Devdas and Jiahua Duojie; Formal Analysis, Jiahua Duojie; Investigation, Samuel Devdas and Jiahua Duojie; Data Curation, Samuel Devdas; Writing --- Original Draft, Jiahua Duojie; Writing --- Review and Editing, Samuel Devdas, Jiahua Duojie, Dr.~Guang Lu, and Dr.~Alexandre de Spindler; Visualization, Jiahua Duojie; Supervision, Dr.~Guang Lu and Dr.~Alexandre de Spindler. All authors have read and agreed to the published version of the manuscript.

\section{Funding}\label{funding}

This research received no external funding.

\section{Institutional Review Board Statement}\label{institutional-review-board-statement}

Not applicable. This study involved no human or animal subjects and used synthetic, non-identifiable, AI-generated dialogues. Contributing authors (Samuel Devdas and Jiahua Duojie) provided qualitative evaluation notes on these synthetic materials; no undisclosed third-party raters and no patient data were involved. Future human subject studies will require full institutional review board approval.

\section{Informed Consent Statement}\label{informed-consent-statement}

Not applicable. No human participants were involved in the simulation study. Future human subject studies will implement comprehensive informed consent protocols explaining AI personality profiling, data use, potential risks, and study procedures.

\section{Acknowledgments}\label{acknowledgments}

The authors acknowledge the PROMISE framework developers {[}7{]} whose open architecture informed our system design. We also acknowledge the prior work in which the D--R--E pipeline, the Evaluation Matrix rubric, and the evaluator methodology used in this study were first conceived {[}68{]}.

\section{Conflicts of Interest}\label{conflicts-of-interest}

The authors declare no conflicts of interest. Qualitative review of evaluation matrices was performed by contributing authors (Samuel Devdas and Jiahua Duojie); no external paid evaluator or undisclosed third-party rater contributed to this study.

